Hydra Battles and Cie

### 2021-08-20 doc/part-hydras.tex
conversion to Alectryon (chapter 4, completed, starting chap. 5)

    \subsubsection{Boundedness}
    If $\alpha$ and $\beta$ are both strictly  less than  $\omega^\gamma$, then so is their natural sum
    $\alpha \oplus \beta$. This result can be proved by structural induction on $\gamma$.

    \begin{Coqsrc}
    Lemma oplus_bounded_phi0 alpha beta gamma :
      nf alpha -> nf beta -> nf gamma ->
      lt alpha (phi0 gamma) ->
      lt beta (phi0 gamma) ->
      lt (alpha o+ beta) (phi0 gamma).
    \end{Coqsrc}

    This lemma helps us

@@

    \begin{Coqsrc}
    Lemma oplus_comm_0 : forall alpha, nf alpha ->
         forall a b,  nf a -> nf b ->
                      lt a alpha ->
                      lt b alpha ->
                      a o+ b = b o+ a.
     Proof with eauto with T1.
        intros alpha Halpha; transfinite_induction alpha.
    (* rest of proof omitted *)  
    \end{Coqsrc}

@@

    \begin{Coqsrc}
      Lemma oplus_comm : forall alpha beta, 
          nf alpha -> nf beta ->
          alpha o+ beta =  beta o+ alpha.
      Proof with eauto with T1.
        intros alpha beta Halpha Hbeta;
        apply oplus_comm_0 with (succ (max alpha beta)) ...  
      (* ... *)
    \end{Coqsrc}

    The associativity of Hessenberg's sum is proved the same way.

    \begin{Coqsrc}
     Lemma oplus_assoc_0 :
        forall alpha,
          nf alpha ->
          forall a b c,  nf a -> nf b -> nf c ->
                          lt a alpha ->
                          lt b alpha -> lt c alpha ->
                          a o+ (b o+ c) = (a o+ b) o+ c.
      Proof with eauto with T1.
        intros alpha Halpha.
        transfinite_induction alpha.
        (* ... *)
    \end{Coqsrc}

    \begin{Coqsrc}
     Lemma oplus_assoc : forall alpha beta gamma,
                            nf alpha -> nf beta -> nf gamma ->
                                        alpha o+ (beta o+ gamma) =
                                        alpha o+ beta o+ gamma.
     Proof with eauto with T1.
        intros;
        apply oplus_assoc_0 with (succ (max alpha (max beta gamma))) ...
        (* ... *)   
    \end{Coqsrc}

@@

    \begin{Coqsrc}
    Lemma oplus_strict_mono_LT_l (alpha beta gamma : T1) :
      nf gamma   -> alpha  t1< beta ->
      alpha o+ gamma  t1< beta o+ gamma.

    Lemma oplus_strict_mono_LT_r (alpha beta gamma : T1) :
      nf alpha -> beta t1< gamma ->
      alpha o+ beta t1< alpha o+ gamma.
    \end{Coqsrc}

@@

    \begin{Coqsrc}
    Fixpoint m (h:Hydra) : T1 :=
      match h with head => zero
                 | node hs => ms hs
    end 
    with ms (s:Hydrae) :  T1 :=
      match s with  hnil => zero
                  | hcons h s' => phi0 (m h) o+  ms s'
     end.  
    \end{Coqsrc}

@@

    \begin{Coqsrc}
    Lemma m_nf : forall h, nf (m h).
    Proof.
     intro h; elim h using Hydra_rect2 
                with (P0 := fun s =>  nf (ms s)).
     (* ... *)

    Lemma ms_nf : forall s, nf (ms s).
    Proof with auto with T1.
    (* ... *)
    \end{Coqsrc}

@@

    \begin{Coqsrc}
    Lemma S0_decr :
      forall s s', S0  s s' -> ms s' t1< ms s.
    \end{Coqsrc}

    \begin{Coqsrc}
    Lemma R1_decr : forall h h',
                      R1 h h' -> m h' t1< m h.
    \end{Coqsrc}

    \begin{Coqsrc}
    Lemma S1_decr n:
      forall s s', S1 n s s' -> ms s' t1<  ms s.
    \end{Coqsrc}

    \begin{Coqsrc}
    Lemma R2_decr n : forall h h', R2 n h h' -> m h'  t1< m h.
    \end{Coqsrc}

    \begin{Coqsrc}
    Lemma round_decr : forall h h', h -1-> h' -> m h' t1< m h.
    Proof.
       destruct 1 as [n [H | H]].
       -  now apply R1_decr.
       -  now apply R2_decr with n.
    Qed.
    \end{Coqsrc}

    Finally, we prove the termination of all (free) battles.

@@

    \begin{Coqsrc}
    Global Instance HVariant : Hvariant lt_wf free var.
    Proof.
     split; intros; eapply round_decr; eauto.
    Qed.

    Theorem every_battle_terminates: Termination.
    Proof. 
      red; apply Inclusion.wf_incl with 
             (R2 := fun h h' =>  m h t1< m h').
       red; intros;  now apply round_decr.
       apply Inverse_Image.wf_inverse_image, T1_wf.
    Qed.
    \end{Coqsrc}

@@

\label{Functions:canonS}
\label{Functions:canon}
\begin{Coqsrc}
  Fixpoint canon alpha (i:nat) : T1 :=
  match alpha with
    zero => zero
  | ocons zero 0 zero  => zero
  | ocons zero (S k) zero  => FS k
  | ocons gamma 0 zero => (match T1.pred gamma with
                            Some gamma' =>
                            match i with 0 => zero
                                    |   S j => ocons gamma' j zero
                            end
                          | None =>
                            ocons (canon gamma i) 0 zero
                           end)

  |  ocons gamma (S n) zero =>
     (match T1.pred gamma with
       Some gamma' =>
       (match i with
         0 =>  ocons gamma n zero
       | S j => ocons gamma n (ocons gamma' j zero)
       end)
       | None =>
            ocons gamma n (ocons (canon gamma  i) 0 zero)
      end)
   |  ocons alpha n beta => ocons alpha n (canon beta i)  
end.
\end{Coqsrc}

@@

\begin{Coqsrc}
Compute pp (canon (T1.omega ^ T1.omega) 3). 
\end{Coqsrc}

\begin{Coqanswer}
  = (_omega ^ 3)%pT1
     : ppT1
   \end{Coqanswer}

\begin{Coqsrc}
Compute pp (canon (T1.omega ^ T1.omega * 3) 42).  
\end{Coqsrc}

\begin{Coqanswer}
    = (_omega ^ _omega * 2 + _omega ^ 42)%pT1
     : ppT1
\end{Coqanswer}

@@

Proving the equality $\canonseq{\alpha+1}{i}=\alpha$ is not 
as simple as suggested by the equations of definition~\ref{def:canonseq-math}\,.
Nevertheless, we can prove it by  plain structural induction on $\alpha$.

\begin{Coqsrc}
  Lemma canon_succ i alpha :
  nf alpha -> canon (succ alpha) i = alpha.
Proof.
 revert i; induction alpha.
 (* ... *)
\end{Coqsrc}

>>>>

@@

    \input{movies/snippets/Hessenberg/oplusComm0}

@@

\input{movies/snippets/Hessenberg/oplusComm}

    Associativity of Hessenberg's sum is proved the same way.

    \input{movies/snippets/Hessenberg/oplusAssoc0}
    \input{movies/snippets/Hessenberg/oplusAssoc}

@@

    \input{movies/snippets/Hessenberg/oplusMono}

@@

   \input{movies/snippets/Hydra_Termination/mDef}

@@

    \input{movies/snippets/Hydra_Termination/mNf}

@@

    \input{movies/snippets/Hydra_Termination/S0Decr}
    \input{movies/snippets/Hydra_Termination/R1Decr}
    \input{movies/snippets/Hydra_Termination/S1Decr}
    \input{movies/snippets/Hydra_Termination/R2Decr}
     \input{movies/snippets/Hydra_Termination/RoundDecr}

      Finally, we prove termination of all (free) battles.

@@

    \input{movies/snippets/Hydra_Termination/FinalThm}

@@

\label{Functions:canonS}
\label{Functions:canon}

\input{movies/snippets/Canon/canonDef}

@@

\input{movies/snippets/Canon/canonExamples}

@@

For instance, the equality $\canonseq{\alpha+1}{i}=\alpha$  can be  proved by  structural induction on $\alpha$.

\input{movies/snippets/Canon/canonSucc}
